Haskell in Computational Biology

### 2013-04-28 jfp.tex
jfp-ified draft

\pagetuningtrue % trying to get page breaks in the best spots

\newif\ifnotcutting \notcuttingfalse % set if not cutting down to size

@@

\documentclass[nonatbib,blockstyle,times]{sigplanconf}
  % at 9pt anything but block style is unreadable

@@

\ifpagetuning
  \newcommand\tuningbox{\vbox}
\else
  \newcomming\tuningbox{\relax}
\fi

\newcommand\naive{na\"\i ve}
\usepackage{url}
\usepackage{amsmath}
\usepackage{array}
\usepackage{listings}
\usepackage{graphicx}
\lstset{
    tabsize = 2,
    basicstyle = \ttfamily,
    language = Haskell
    }

@@

\newif\ifverbatimsmall
\verbatimsmallfalse
\makeatletter
\def\verbatim@font{%
   \normalfont\ttfamily
   \ifverbatimsmall\small\fi
   \hyphenchar\font\m@ne
   \@noligs}
\makeatother

@@

\begin{document}
%\parskip=0.9\parskip % cheating
%\advance \parskip by 0pt plus 0.1pt

@@

\conferenceinfo{ICFP'12,} {September 9--15, 2012, Copenhagen, Denmark.}
\CopyrightYear{2012}
\copyrightdata{978-1-4503-1054-3/12/09}

%\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{Haskell in Computational Biology}   % 'preprint' option specified.
\iftimestamp
\preprintfooter{\mdfivestamp} 
\fi

\title{Experience Report: Haskell in Computational Biology}
% \subtitle{Subtitle Text, if any}

\authorinfo{Noah M. Daniels \and Andrew Gallant \and Norman Ramsey}
           {Department of Computer Science, Tufts University}
           {{\rmfamily\{}ndaniels, agallant, nr{\rmfamily\}}@cs.tufts.edu}

@@

\begingroup
\ifpagetuning
\makeatletter
\@paragraphaboveskip = 0.5\@paragraphaboveskip
\makeatother
\fi

\category{D.1.1}{Applicative (Functional) Programming}{}
\category{J.3}{Biology and genetics}{}
% \category{CR-number}{subcategory}{third-level}
% 
% \terms
% term1, term2
% 
\keywords
memoization, stochastic search, parallel strategies,  QuickCheck,
remote homology detection 
{\raggedright\par}

\endgroup

@@

\acks

Anonymous referees spurred us to think
more deeply about laziness and to write more carefully about performance.
Josef Svenningsson suggested a correspondence between \mrfy's search
and stream fusion, which led us eventually to the \texttt{Utility} type.
Koen Claessen instantly diagnosed an inappropriately strict
implementation of the \texttt{Rand} monad.
We also thank Lenore Cowen, Kathleen Fisher, Ben Hescott, Brad
Larsen, and Nathan Ricci.

@@

\bibliographystyle{plainnat}

\iffalse
\bibliography{mrfy_experience_report}
\else
\begin{thebibliography}{9}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Cowen et~al.(2002)Cowen, Bradley, Menke, King, and
  Berger]{Cowen:2002p588}
Lenore Cowen, Philip Bradley, Matt Menke, Jonathan King, and Bonnie Berger.
\newblock {Predicting the beta-helix fold from protein sequence data}.
\newblock \emph{Journal of Computational Biology}, 2002.

\bibitem[Daniels et~al.(2012)Daniels, Hosur, Berger, and Cowen]{Daniels:2012}
Noah Daniels, Raghavendra Hosur, Bonnie Berger, and Lenore Cowen.
\newblock {SMURFLite: combining simplified Markov random fields with simulated
  evolution improves remote homology detection for beta-structural proteins
  into the twilight zone}.
\newblock \emph{Bioinformatics}, March 2012.

\bibitem[Durbin et~al.(1998)Durbin, Eddy, Krogh, and Mitchison]{Durbin:1998wz}
Rirchard Durbin, Sean Eddy, Anders Krogh, and Graeme Mitchison.
\newblock \emph{{Biological Sequence Analysis: Probabilistic Models of Proteins
  and Nucleic Acids}}.
\newblock Cambridge University Press, May 1998.

\bibitem[Eddy(1998)]{Eddy:1998ut}
Sean Eddy.
\newblock {Profile hidden Markov models}.
\newblock \emph{Bioinformatics}, 14:\penalty0 755--763, 1998.

\bibitem[Hughes(1989)]{hughes:why}
John Hughes.
\newblock Why functional programming matters.
\newblock \emph{The Computer Journal}, 32\penalty0 (2):\penalty0 98--107, April
  1989.

\bibitem[Menke et~al.(2008)Menke, Berger, and Cowen]{Menke:2008wu}
Matthew Menke, Bonnie Berger, and Lenore Cowen.
\newblock {Matt: local flexibility aids protein multiple structure alignment}.
\newblock \emph{PLoS Computational Biology}, 2008.

\bibitem[Menke et~al.(2010)Menke, Berger, and Cowen]{Menke:2010ti}
Matthew Menke, Bonnie Berger, and Lenore Cowen.
\newblock {Markov random fields reveal an N-terminal double beta-propeller
  motif as part of a bacterial hybrid two-component sensor system}.
\newblock \emph{Proceedings of the National Academy of Science}, 2010.

\bibitem[Sansom and Peyton~Jones(1997)]{sansom-pj}
Patrick Sansom and Simon~L Peyton~Jones.
\newblock Formally based profiling for higher-order functional languages.
\newblock \emph{ACM TOPLAS}, 19\penalty0 (2):\penalty0 334--385, 1997.

\bibitem[Viterbi(1967)]{Viterbi:1967hq}
Andrew Viterbi.
\newblock {Error bounds for convolutional codes and an asymptotically optimum
  decoding algorithm}.
\newblock \emph{IEEE Transactions on Information Theory}, 13\penalty0
  (2):\penalty0 260--269, April 1967.

\end{thebibliography}
\fi

\iffinaldraft

\vfill

\begingroup
\parfillskip=0pt
\advance\leftskip by 0pt plus 30em
\emph{You can live to surf the Haskell wave, but~if you slide off the crest, you
drown.}
\par
\endgroup

\fi

>>>>

\pagetuningfalse % trying to get page breaks in the best spots

\newif\ifnotcutting \notcuttingtrue % set if not cutting down to size

@@

\documentclass[compmod]{jfp1}

@@

\ifpagetuning
  \newcommand\tuningbox{\vbox}
\else
  \newcommand\tuningbox{\relax}
\fi

\newcommand\naive{na\"\i ve}
\usepackage{url}
%\usepackage{amsmath}
\usepackage{array}
\usepackage{listings}
\usepackage{graphicx}
\lstset{
    tabsize = 2,
    basicstyle = \ttfamily,
    language = Haskell
    }

@@

\newif\ifverbatimsmall
\verbatimsmallfalse
\makeatletter
\def\verbatim@font{%
   \normalfont\ttfamily
   \ifverbatimsmall\small\fi
   \hyphenchar\font\m@ne
   \@noligs}
\newcommand{\mono}[1]{%
  {\@tempdima = \fontdimen2\font
   \texttt{\spaceskip = 1.1\@tempdima #1}}}
\makeatother

@@

\title{Haskell in Computational Biology}

\author[N. M. Daniels, A. Gallant, and N. Ramsey]
        {Noah M. Daniels \and Andrew Gallant \and Norman Ramsey \\
           {Department of Computer Science, Tufts University}\\
           {\email{{\rmfamily\{}ndaniels, agallant, nr{\rmfamily\}}@cs.tufts.edu}}}

\begin{document}

@@

\iftimestamp
%\preprintfooter{\mdfivestamp} 
\fi

@@

@@

\section*{Acknowledgments}

Anonymous referees spurred us to think
more deeply about laziness and to write more carefully about performance.
Josef Svenningsson suggested a correspondence between \mrfy's search
and stream fusion, which led us eventually to the \texttt{Utility} type.
Koen Claessen instantly diagnosed an inappropriately strict
implementation of the \texttt{Rand} monad.
John Tibell taught us the \texttt{UNPACK} pragma.
We also thank Lenore Cowen, Kathleen Fisher, Ben Hescott, Brad
Larsen, and Nathan Ricci.

@@

\bibliography{mrfy_experience_report}
\bibliographystyle{jfp}

%%  \iffinaldraft
%%  
%%  
%%  \vfill
%%  
%%  \begingroup
%%  \parfillskip=0pt
%%  \advance\leftskip by 0pt plus 30em
%%  \emph{You can live to surf the Haskell wave, but~if you slide off the crest, you
%%  drown.}
%%  \par
%%  \endgroup
%%  
%%  \fi
